### 2023-02-11 DraftDataAnalysis.Rmd
Add files via upload

---
title: "DraftDataAnalysis"
author: "Kyle McCullough"
date: "2023-02-11"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```