### 2023-03-03 DraftDataAnalysis.Rmd
Add files via upload

---
title: "DraftDataAnalysis"
author: "Kyle McCullough"
date: "2023-02-11"
output: word_document
---

#Loading cleaned draft data
```{r}
library(tidyverse)
library(readxl)

draft = read_excel('1993-2022draftStats.xlsx', col_names = TRUE)
draft

draft$wAV[is.na(draft$wAV)] = 0
draft$DrAV[is.na(draft$DrAV)] = 0
draft$G[is.na(draft$G)] = 0
draft$Cmp[is.na(draft$Cmp)] = 0
draft$PassAtt[is.na(draft$PassAtt)] = 0
draft$PassYds[is.na(draft$PassYds)] = 0
draft$PassTD[is.na(draft$PassTD)] = 0
draft$`Int(O)`[is.na(draft$`Int(O)`)] = 0
draft$RushAtt[is.na(draft$RushAtt)] = 0
draft$RushYds[is.na(draft$RushYds)] = 0
draft$RushTD[is.na(draft$RushTD)] = 0
draft$Rec[is.na(draft$Rec)] = 0
draft$RecYds[is.na(draft$RecYds)] = 0
draft$RecTD[is.na(draft$RecTD)] = 0
draft$Solo[is.na(draft$Solo)] = 0
draft$`Int(D)`[is.na(draft$`Int(D)`)] = 0
draft$Sk[is.na(draft$Sk)] = 0
draft$`College/Univ`[3051] = "Bethel (TN)"
draft$`College/Univ`[3145] = "Lousiana-Lafayette"
draft$`College/Univ`[3149] = "Regina"
draft$`College/Univ`[3170] = "Lousiana-Lafayette"
draft$`College/Univ`[3229] = "California Polytechnic State"
draft$`College/Univ`[3231] = "Missouri Western State"
draft$`College/Univ`[3249] = "Georgia State"
draft$`College/Univ`[3261] = "Alabama-Birmingham"
draft$`College/Univ`[3516] = "Colorado State-Pueblo"
draft$`College/Univ`[3555] = "Northeastern State"
draft$`College/Univ`[4259] = "No college"
draft$`College/Univ`[4818] = "No college"
draft$`College/Univ`[6046] = "Eastern Michigan"
draft$`College/Univ`[6843] = "Trinity International"
draft$`College/Univ`[7097] = "Trinity International"
draft$`College/Univ`[7296] = "Clark Atlanta"

this = draft %>%
  mutate(Fran = Tm)

this = this %>%
  mutate(Fran = replace(Fran, Fran == "PHO", "ARI")) %>%
  mutate(Fran = replace(Fran, Fran == "SDG", "LAC")) %>%
  mutate(Fran = replace(Fran, Fran == "STL", "LAR")) %>%
  mutate(Fran = replace(Fran, Fran == "RAM", "LAR")) %>%
  mutate(Fran = replace(Fran, Fran == "OAK", "LVR")) %>%
  mutate(Fran = replace(Fran, Fran == "RAI", "LVR"))
```

Pro Football Reference has provided a wAV stat which is the weigthed average value
of each player and this is what I will be using to show how productive of a career
a player had.
```{r}
draftPick = draft %>%
  filter(Draft < 2011)

ggplot(draftPick, aes(Pick, wAV)) +
  geom_point()
reg = lm(draft$wAV~draft$Pick)
jtools :: summ(reg)
anova(reg)
```

Which teams have had the most successful draft picks?
```{r}
teamDraftSuccess = this %>%
  group_by(Fran) %>%
  summarise(averagewAV = mean(wAV))
teamDraftSuccess = arrange(teamDraftSuccess, desc(averagewAV))
teamDraftSuccess
```