### 2023-05-07 DraftDataAnalysis.Rmd
Add files via upload

---
title: "DraftDataAnalysis"
author: "Kyle McCullough"
date: "2023-02-11"
output: word_document
---

#Loading cleaned draft data
```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(nflplotR)

draft = read_excel('1993-2022draftStats.xlsx', col_names = TRUE)
draft

draft$wAV[is.na(draft$wAV)] = 0
draft$DrAV[is.na(draft$DrAV)] = 0
draft$G[is.na(draft$G)] = 0
draft$Cmp[is.na(draft$Cmp)] = 0
draft$PassAtt[is.na(draft$PassAtt)] = 0
draft$PassYds[is.na(draft$PassYds)] = 0
draft$PassTD[is.na(draft$PassTD)] = 0
draft$`Int(O)`[is.na(draft$`Int(O)`)] = 0
draft$RushAtt[is.na(draft$RushAtt)] = 0
draft$RushYds[is.na(draft$RushYds)] = 0
draft$RushTD[is.na(draft$RushTD)] = 0
draft$Rec[is.na(draft$Rec)] = 0
draft$RecYds[is.na(draft$RecYds)] = 0
draft$RecTD[is.na(draft$RecTD)] = 0
draft$Solo[is.na(draft$Solo)] = 0
draft$`Int(D)`[is.na(draft$`Int(D)`)] = 0
draft$Sk[is.na(draft$Sk)] = 0
draft$`College/Univ`[3051] = "Bethel (TN)"
draft$`College/Univ`[3145] = "Lousiana-Lafayette"
draft$`College/Univ`[3149] = "Regina"
draft$`College/Univ`[3170] = "Lousiana-Lafayette"
draft$`College/Univ`[3229] = "California Polytechnic State"
draft$`College/Univ`[3231] = "Missouri Western State"
draft$`College/Univ`[3249] = "Georgia State"
draft$`College/Univ`[3261] = "Alabama-Birmingham"
draft$`College/Univ`[3516] = "Colorado State-Pueblo"
draft$`College/Univ`[3555] = "Northeastern State"
draft$`College/Univ`[4259] = "No college"
draft$`College/Univ`[4818] = "No college"
draft$`College/Univ`[6046] = "Eastern Michigan"
draft$`College/Univ`[6843] = "Trinity International"
draft$`College/Univ`[7097] = "Trinity International"
draft$`College/Univ`[7296] = "Clark Atlanta"

this = draft %>%
  mutate(Fran = Tm)

this = this %>%
  mutate(Fran = replace(Fran, Fran == "PHO", "ARI")) %>%
  mutate(Fran = replace(Fran, Fran == "SDG", "LAC")) %>%
  mutate(Fran = replace(Fran, Fran == "STL", "LAR")) %>%
  mutate(Fran = replace(Fran, Fran == "RAM", "LAR")) %>%
  mutate(Fran = replace(Fran, Fran == "OAK", "LVR")) %>%
  mutate(Fran = replace(Fran, Fran == "RAI", "LVR"))

#draft = draft %>%
  #mutate(Pos = replace(Pos, Pos == "S", "DB")) %>%
  #mutate(Pos = replace(Pos, Pos == "CB", "DB")) %>%
  #mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  #mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  #mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  #mutate(Pos = replace(Pos, Pos == "OLB", "LB")) %>%
  #mutate(Pos = replace(Pos, Pos == "ILB", "LB"))
```
```{r}
draft %>%
  filter(Pos == c("K")) %>%
  #filter(wAV <= 9) %>%
  filter(wAV >= 35) %>%
  summarise(averageAP1 = mean(AP1), averagePB = mean(PB), averageSt = mean(St), averageG = mean(G), averageCompPer = mean(Cmp/PassAtt), averagePassYds = mean(PassYds), averagePassTds = mean(PassTD), averageIntO = mean(`Int(O)`), averageRushYds = mean(RushYds), averageRushTds = mean(RushTD), averageRec = mean(Rec), averageRecYds = mean(RecYds), averageRecTds = mean(RecTD), averageSolo = mean(Solo), averageIntD = mean(`Int(D)`), averageSk = mean(Sk))
```

Pro Football Reference has provided a wAV stat which is the weigthed average value
of each player and this is what I will be using to show how productive of a career
a player had.
```{r}
draftPick = draft %>%
  filter(Draft < 2011)

reg1 = lm(draftPick$wAV~draftPick$Pick)
reg2 = lm(draftPick$wAV~log(draftPick$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

draftPick %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              ) +
  labs(title = "Average value of each draft pick",
       subtitle = "Data from 1993 to 2010")

plot.fitted <- function(lmod) {
  require(ggplot2,dplyr)
  df = lmod$model
  title = paste("Residuals vs. Fits (Dependent Variable: ",
                names(df)[1],
                ")",
                sep = "")
  pr <- df %>% ggplot(aes(x = fitted(lmod), y = resid(lmod))) +
    geom_ribbon(aes(ymin = -sigma(lmod), ymax = sigma(lmod)),
                fill = "gray",
                alpha = .5) +
    geom_ribbon(aes(ymin = -sigma(lmod) * 2, ymax = sigma(lmod) * 2),
                fill = "lightgray",
                alpha = .5) +
    geom_point() +
    theme_classic() +
    ggtitle(title) +
    ylab("Residules") +
    xlab("Fitted") +
    geom_hline(yintercept = 0,
               linetype = "solid",
               color = "black")
  return(pr)  
}

plot.fitted(reg1)
plot.fitted(reg2)

b = reg2$fitted.values
PickValue = tibble(Pick = 1:4489, wAV = b)
PickValue = PickValue %>%
  filter(Pick < 256)
PickValue
```

```{r}
roundPick = draftPick %>%
  filter(Draft > 1993) %>%
  group_by(Rnd) %>%
  summarise(averagewAV = mean(wAV))
roundPick
```
```{r}
draftPick %>%
  group_by(Pick) %>%
  summarise(averagewAV = mean(wAV))
```

Which teams have had the most successful draft picks?
```{r}
teamDraftSuccess = this %>%
  group_by(Fran) %>%
  summarise(averagewAV = mean(wAV))
teamDraftSuccess = arrange(teamDraftSuccess, desc(averagewAV))
teamDraftSuccess

teamDraftSuccess %>%
  ggplot2::ggplot(aes(reorder(Fran, averagewAV), averagewAV)) +
  nflplotR::geom_nfl_logos(aes(team_abbr = Fran), width = 0.065, alpha = 0.7) +
  ggplot2::labs(title = "Average Value of Player Drafted by each Franchise",
       subtitle = "From 1993 to 2022",
       y = "Average Value per Player drafted",
       x = "Franchise") +
  ggplot2::theme(plot.title = ggplot2::element_text(face = "bold"),
                 axis.text.x = element_blank(),
                 axis.ticks.x = element_blank())
```

Which college has the best nfl prospects?
```{r}
draft %>%
  group_by(`College/Univ`) %>%
  summarise(sumwAV = sum(wAV)) %>%
  arrange(desc(sumwAV))

draft %>%
  filter(Draft < 2011) %>%
  group_by(`College/Univ`) %>%
  filter(n() > 19) %>%
  summarise(averagewAV = mean(wAV)) %>%
  arrange(desc(averagewAV)) 
```

```{r}
draft %>%
  filter(Pos == "WR") %>%
  group_by(`College/Univ`) %>%
  summarise(sumwAV = sum(wAV)) %>%
  arrange(desc(sumwAV))

draft %>%
  filter(Pos == "WR") %>%
  filter(`College/Univ` == "Ohio St.") %>%
  arrange(desc(wAV))

draft %>%
  filter(Pos == "WR") %>%
  filter(`College/Univ` == "LSU") %>%
  arrange(desc(wAV))
```
```{r}
draft %>%
  filter(Pos == "DB") %>%
  group_by(`College/Univ`) %>%
  summarise(sumwAV = sum(wAV)) %>%
  arrange(desc(sumwAV))

draft %>%
  filter(Pos == "DB") %>%
  filter(`College/Univ` == "Alabama") %>%
  arrange(desc(wAV))

draft %>%
  filter(Pos == "DB") %>%
  filter(`College/Univ` == "LSU") %>%
  arrange(desc(wAV))
```

```{r}
draft %>%
  filter(`College/Univ` == "Boston Col.") %>%
  filter(Draft < 2011) %>%
  arrange(desc(wAV))

draft %>%
  filter(`College/Univ` == "Miami (FL)") %>%
  filter(Draft < 2011) %>%
  arrange(desc(wAV))
```

```{r}
draft %>%
  filter(Pos == c("K"))%>%
  arrange(desc(wAV))
draft %>%
  filter(Player == "Cameron Heyward")
```

Which Positions have the most value?
```{r}
draft %>%
  group_by(Pos) %>%
  summarise(averagewAV = mean(wAV)) %>%
  arrange(desc(averagewAV))
```
Which was the best NFL draft?
```{r}
draft %>%
  group_by(Draft) %>%
  summarize(sumwAV = sum(wAV)) %>%
  arrange(desc(sumwAV)) 
draft %>%
  filter(Draft == 2009) %>%
  arrange(desc(wAV))
draft %>%
  filter(Pos == "WR") %>%
  filter(wAV == 4)
```
When do positions tend to get drafted?
```{r}
draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  group_by(Pos) %>%
  summarise(avgDraftPos = mean(Pick)) %>%
  arrange(avgDraftPos) 
```
When the first player at each position gets picked on average?
```{r}
draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  arrange(Pick) %>%
  group_by(Pos, Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft) %>%
  filter(PosDraft == 1) %>%
  summarise(avgPick = mean(Pick)) %>%
  arrange(avgPick)
```

When should positions really be drafted?
AveragewAV for first round talent per Position
```{r}
draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Draft < 2011) %>%
  filter(Rnd == 1) %>%
  group_by(Pos) %>%
  summarize(avgwAV = mean(wAV)) %>%
  arrange(desc(avgwAV))
```

Position: T
wAV as the Draft goes on and then the wAV per each player per position in draft
```{r}
draftT = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "T") %>%
  filter(Draft < 2011) %>% 
  arrange(Pick)

attach(draftT) 
reg1 = lm(draftT$wAV~draftT$Pick)
reg2 = lm(draftT$wAV~log(draftT$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

#wAV per Pick per tackle
p = draftT %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftTOrder = draftT %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftTOrder) 
reg1 = lm(draftTOrder$wAV~draftTOrder$PosDraft)
reg2 = lm(draftTOrder$wAV~log(draftTOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftTOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: QB
```{r}
draftQB = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "QB") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftQB) 
reg1 = lm(draftQB$wAV~draftQB$Pick)
reg2 = lm(draftQB$wAV~log(draftQB$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftQB %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftQBOrder = draftQB %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftQBOrder) 
reg1 = lm(draftQBOrder$wAV~draftQBOrder$PosDraft)
reg2 = lm(draftQBOrder$wAV~log(draftQBOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftQBOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: DE
```{r}
draftDE = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "DE") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftDE) 
reg1 = lm(draftDE$wAV~draftDE$Pick)
reg2 = lm(draftDE$wAV~log(draftDE$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftDE %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftDEOrder = draftDE %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftDEOrder) 
reg1 = lm(draftDEOrder$wAV~draftDEOrder$PosDraft)
reg2 = lm(draftDEOrder$wAV~log(draftDEOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftDEOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: DT
```{r}
draftDT = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "DT") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftDT) 
reg1 = lm(draftDT$wAV~draftDT$Pick)
reg2 = lm(draftDT$wAV~log(draftDT$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftDT %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftDTOrder = draftDT %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftDTOrder) 
reg1 = lm(draftDTOrder$wAV~draftDTOrder$PosDraft)
reg2 = lm(draftDTOrder$wAV~log(draftDTOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftDTOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: C
```{r}
draftC = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "C") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftC) 
reg1 = lm(draftC$wAV~draftC$Pick)
reg2 = lm(draftC$wAV~log(draftC$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftC %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftCOrder = draftC %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftCOrder) 
reg1 = lm(draftCOrder$wAV~draftCOrder$PosDraft)
reg2 = lm(draftCOrder$wAV~log(draftCOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftCOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: G
```{r}
draftG = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "G") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftG) 
reg1 = lm(draftG$wAV~draftG$Pick)
reg2 = lm(draftG$wAV~log(draftG$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftG %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftGOrder = draftG %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftGOrder) 
reg1 = lm(draftGOrder$wAV~draftGOrder$PosDraft)
reg2 = lm(draftGOrder$wAV~log(draftGOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftGOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: LB
```{r}
draftLB = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "LB") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftLB) 
reg1 = lm(draftLB$wAV~draftLB$Pick)
reg2 = lm(draftLB$wAV~log(draftLB$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftLB %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftLBOrder = draftLB %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftLBOrder) 
reg1 = lm(draftLBOrder$wAV~draftLBOrder$PosDraft)
reg2 = lm(draftLBOrder$wAV~log(draftLBOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftLBOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: RB
```{r}
draftRB = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "RB") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftRB) 
reg1 = lm(draftRB$wAV~draftRB$Pick)
reg2 = lm(draftRB$wAV~log(draftRB$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftRB %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftRBOrder = draftRB %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftRBOrder) 
reg1 = lm(draftRBOrder$wAV~draftRBOrder$PosDraft)
reg2 = lm(draftRBOrder$wAV~log(draftRBOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftRBOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: DB
```{r}
draftDB = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "DB") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftDB) 
reg1 = lm(draftDB$wAV~draftDB$Pick)
reg2 = lm(draftDB$wAV~log(draftDB$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftDB %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftDBOrder = draftDB %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftDBOrder) 
reg1 = lm(draftDBOrder$wAV~draftDBOrder$PosDraft)
reg2 = lm(draftDBOrder$wAV~log(draftDBOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftDBOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: WR
```{r}
draftWR = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "WR") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftWR) 
reg1 = lm(draftWR$wAV~draftWR$Pick)
reg2 = lm(draftWR$wAV~log(draftWR$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftWR %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftWROrder = draftWR %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftWROrder) 
reg1 = lm(draftWROrder$wAV~draftWROrder$PosDraft)
reg2 = lm(draftWROrder$wAV~log(draftWROrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftWROrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```
Position: TE
```{r}
draftTE = draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Pos == "TE") %>%
  filter(Draft < 2011) %>%
  arrange(Pick)

attach(draftTE) 
reg1 = lm(draftTE$wAV~draftTE$Pick)
reg2 = lm(draftTE$wAV~log(draftTE$Pick))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p = draftTE %>% ggplot(aes(x = Pick, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p
p2 = ggplot_build(p)
p2$data[[2]]

draftTEOrder = draftTE %>%
  group_by(Draft) %>%
  mutate(NumPosInDraft = n()) %>%
  mutate(PosDraft = 1:NumPosInDraft) %>%
  arrange(PosDraft) %>%
  ungroup(Draft)

attach(draftTEOrder) 
reg1 = lm(draftTEOrder$wAV~draftTEOrder$PosDraft)
reg2 = lm(draftTEOrder$wAV~log(draftTEOrder$PosDraft))
jtools::export_summs(reg1, reg2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

p3 = draftTEOrder %>%
  ggplot(aes(x=PosDraft, y=wAV)) +
  geom_point() +
  stat_smooth(method = lm,
              formula = y~log(x),
              se = TRUE,
              )
p3
p4 = ggplot_build(p3)
p4$data[[2]]
```

```{r}
draft %>%
  mutate(Pos = replace(Pos, Pos == "SAF", "DB")) %>%
  mutate(Pos = replace(Pos, Pos == "NT", "DT")) %>%
  mutate(Pos = replace(Pos, Pos == "OT", "T")) %>%
  filter(Draft < 2011) %>%
  group_by(Pos) %>%
  summarize(avgwAV = mean(wAV)) %>%
  arrange(desc(avgwAV))

draft %>%
  filter(Draft < 2011) %>%
  group_by(Pos) %>%
  summarize(avgwAV = mean(wAV))
draft %>%
  filter(Pos == "DL")
```