### 2023-05-07 DraftPlayerPrediction.Rmd

---
title: "Untitled"
output: word_document
date: "2023-04-21"
---

#load cleaned data
```{r}
library(tidyverse)
library(readxl)
library(dplyr)

draft = read_excel('1993-2022draftStats.xlsx', col_names = TRUE)
draft

draft$wAV[is.na(draft$wAV)] = 0
draft$DrAV[is.na(draft$DrAV)] = 0
draft$G[is.na(draft$G)] = 0
draft$Cmp[is.na(draft$Cmp)] = 0
draft$PassAtt[is.na(draft$PassAtt)] = 0
draft$PassYds[is.na(draft$PassYds)] = 0
draft$PassTD[is.na(draft$PassTD)] = 0
draft$`Int(O)`[is.na(draft$`Int(O)`)] = 0
draft$RushAtt[is.na(draft$RushAtt)] = 0
draft$RushYds[is.na(draft$RushYds)] = 0
draft$RushTD[is.na(draft$RushTD)] = 0
draft$Rec[is.na(draft$Rec)] = 0
draft$RecYds[is.na(draft$RecYds)] = 0
draft$RecTD[is.na(draft$RecTD)] = 0
draft$Solo[is.na(draft$Solo)] = 0
draft$`Int(D)`[is.na(draft$`Int(D)`)] = 0
draft$Sk[is.na(draft$Sk)] = 0
draft$`College/Univ`[3051] = "Bethel (TN)"
draft$`College/Univ`[3145] = "Lousiana-Lafayette"
draft$`College/Univ`[3149] = "Regina"
draft$`College/Univ`[3170] = "Lousiana-Lafayette"
draft$`College/Univ`[3229] = "California Polytechnic State"
draft$`College/Univ`[3231] = "Missouri Western State"
draft$`College/Univ`[3249] = "Georgia State"
draft$`College/Univ`[3261] = "Alabama-Birmingham"
draft$`College/Univ`[3516] = "Colorado State-Pueblo"
draft$`College/Univ`[3555] = "Northeastern State"
draft$`College/Univ`[4259] = "No college"
draft$`College/Univ`[4818] = "No college"
draft$`College/Univ`[6046] = "Eastern Michigan"
draft$`College/Univ`[6843] = "Trinity International"
draft$`College/Univ`[7097] = "Trinity International"
draft$`College/Univ`[7296] = "Clark Atlanta"

combine = read_excel('1993-2022combineData.xlsx', col_names = TRUE)
combine

colnames(combine)[1] <- "Draft"
colnames(combine)[2] <- "Player"
colnames(combine)[4] <- "Pos"
combine = combine %>%
  drop_na(Player)
combine

combinePast = combine %>%
  filter(Draft < 2023)

combinePast = combinePast[,-3]
combinePast = combinePast[,-3]

draftCombine = merge(x = draft, y = combinePast, by = c("Player", "Draft"), all.x = TRUE)
draftCombine
```
```{r}
bigChartQB %>%
  group_by(Conf) %>%
  summarise(meanwAV = mean(wAV)) %>%
  arrange(desc(meanwAV))

bigChartQB %>%
  group_by(level) %>%
  summarise(meanwAV = mean(wAV)) %>%
  arrange(desc(meanwAV))
```

```{r}
bigChartQB %>% ggplot(aes(x = `Weight (lbs)`, y = wAV)) +
  geom_point() + 
  stat_smooth(method = lm,
              formula = y~x^2,
              se = TRUE,
              )

bigChartQB %>%
  filter(complete.cases(`Height (in)`)) %>%
  ggplot(aes(x = `Height (in)`, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(`Weight (lbs)`)) %>%
  ggplot(aes(x = `Weight (lbs)`, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(`Hand Size (in)`)) %>%
  ggplot(aes(x = `Hand Size (in)`, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(`Arm Length (in)`)) %>%
  ggplot(aes(x = `Arm Length (in)`, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(Wonderlic)) %>%
  ggplot(aes(x = Wonderlic, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(`40 Yard`)) %>%
  ggplot(aes(x = `40 Yard`, y = wAV)) +
  geom_point()

bigChartQB %>%
  filter(complete.cases(`Vert Leap (in)`)) %>%
  ggplot(aes(x = `Vert Leap (in)`, y = wAV)) +
  geom_point()

```
```{r}
mod1 = lm(bigChartQB$wAV~I(bigChartQB$`Arm Length (in)`^2))
mod2 = lm(bigChartQB$wAV~bigChartQB$CollegeRAvg)
jtools::export_summs(mod1, mod2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))
```

```{r}
#all sig left
regQB1 = lm(bigChartQB$wAV~bigChartQB$`Hand Size (in)` + bigChartQB$CollegeRYds + bigChartQB$CollegeRAvg)
#all
regQB2 = lm(bigChartQB$wAV ~ bigChartQB$`Height (in)` + bigChartQB$`Hand Size (in)` + bigChartQB$`Weight (lbs)` + bigChartQB$`Arm Length (in)`
            + bigChartQB$`40 Yard` + bigChartQB$`Vert Leap (in)` + bigChartQB$`Broad Jump (in)` + bigChartQB$Shuttle + bigChartQB$`3Cone`
            + bigChartQB$CollegeG + bigChartQB$CollegeAtt + bigChartQB$CollegeCmp + bigChartQB$CollegePct 
            + bigChartQB$CollegeRYds + bigChartQB$CollegeYds + bigChartQB$`CollegeAY/A` + bigChartQB$`CollegeY/A` + bigChartQB$CollegeTD
            + bigChartQB$CollegeInt + bigChartQB$CollegeRTD + bigChartQB$CollegeRate + bigChartQB$CollegeRAtt + bigChartQB$CollegeRAvg)
#one by one
regQB3 = lm(bigChartQB$wAV~bigChartQB$CollegeRYds)
#one by one sig all
regQB4 = lm(bigChartQB$wAV~bigChartQB$`Height (in)` + bigChartQB$`Weight (lbs)` + bigChartQB$`Hand Size (in)` + bigChartQB$`Vert Leap (in)` 
            + bigChartQB$CollegePct)
#one by one sig all left
regQB5 = lm(bigChartQB$wAV~bigChartQB$`Hand Size (in)`)
#one by one sig all each left
regQB6 = lm(bigChartQB$wAV~bigChartQB$`Height (in)` + bigChartQB$CollegePct)
#sig values: Height3, Weight2, HandSize2, Wonderlic2, Vertical Leap1, CollegePct1
#not sig values: armlength, 40yard, BroadJump, Shuttle, 3Cone, CollegeG, CollegeCmp, CollegeAtt, CollegeYds, CollegeTD, CollegeAYA, CollegeYA, Int, Rate, Rush
jtools::export_summs(regQB1, regQB2, regQB3, regQB4, regQB5, regQB6, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))
```
```{r}
Power5QB = bigChartQB %>%
  filter(level == "Power")

modP1 = lm(Power5QB$wAV~I(Power5QB$CollegeRAvg^2))
modP2 = lm(Power5QB$wAV~Power5QB$CollegeRAvg)
jtools::export_summs(modP1, modP2, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))

#sig values: Height2, CollegePct1, 
regPQB = lm(Power5QB$wAV~I(Power5QB$`Height (in)`^2) + Power5QB$CollegePct)
regPQBallquad = lm(Power5QB$wAV~I(Power5QB$`Height (in)`^2) + Power5QB$`Hand Size (in)` + Power5QB$`Weight (lbs)` + Power5QB$`Arm Length (in)`
                   + I(Power5QB$`40 Yard`^2) + Power5QB$`Vert Leap (in)` + Power5QB$`Broad Jump (in)` + Power5QB$Shuttle + Power5QB$`3Cone`
                   + Power5QB$CollegeG + I(Power5QB$CollegeAtt^2) + I(Power5QB$CollegeCmp^2) + Power5QB$CollegePct + I(Power5QB$CollegeRYds^2) 
                   + I(Power5QB$CollegeYds^2) + Power5QB$`CollegeAY/A` + Power5QB$`CollegeY/A` + Power5QB$CollegeTD + Power5QB$CollegeInt
                   + Power5QB$CollegeRTD + Power5QB$CollegeRate + Power5QB$CollegeRAtt + Power5QB$CollegeRAvg)
regPQBall = lm(Power5QB$wAV ~ Power5QB$`Height (in)` + Power5QB$`Hand Size (in)` + Power5QB$`Weight (lbs)` + Power5QB$`Arm Length (in)`
            + Power5QB$`40 Yard` + Power5QB$`Vert Leap (in)` + Power5QB$`Broad Jump (in)` + Power5QB$Shuttle + Power5QB$`3Cone`
            + Power5QB$CollegeG + Power5QB$CollegeAtt + Power5QB$CollegeCmp + Power5QB$CollegePct 
            + Power5QB$CollegeRYds + Power5QB$CollegeYds + Power5QB$`CollegeAY/A` + Power5QB$`CollegeY/A` + Power5QB$CollegeTD
            + Power5QB$CollegeInt + Power5QB$CollegeRTD + Power5QB$CollegeRate + Power5QB$CollegeRAtt + Power5QB$CollegeRAvg)
regPQBallsig = lm(Power5QB$wAV ~ Power5QB$`Height (in)` + Power5QB$`Hand Size (in)`+ Power5QB$`40 Yard` 
            + Power5QB$CollegeYds + Power5QB$CollegeRAvg)
jtools::export_summs(regPQB, regPQBallquad, regPQBall, regPQBallsig, digits=3, statistics = c(N = "nobs",R2 = "r.squared",AIC = "AIC",BIC = "BIC"))
```